\bfseries\LARGE\singlespacing{Transmission of airborne respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-07-10 doc/paper/manuscript.tex
Comparison with previous study

In a previous study\cite{Banholzer2023PLoSMed}, we estimated the effectiveness of mask wearing and air cleaners during the SARS-CoV-2 Omicron wave. \Cref{fig:comparison} shows a comparison of the results. The distribution of respiratory infections is different (\Cref{fig:comparison}a) because we almost exclusively detected SARS-CoV-2 in our previous study during the COVID-19 pandemic, as compared to a range of respiratory infections in this study during non-pandemic times. Fewer respiratory infections were detected in bioaerosols and on the HEPA-filters of the air cleaners in this study (\Cref{fig:comparison}b).
The reductions in particle concentrations were similar (\Cref{fig:comparison}c). The risk of infection tended to be lower with air cleaners in this study where air cleaners were analyzed using a cross-over design, but not in our previous study where air cleaners were installed only at the end of the study when most students had already been infected with SARS-Cov-2(\Cref{fig:comparison}d).

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/epi-data/cases_by_week.pdf} 
    \includegraphics[width=\linewidth]{../../results/epi-data/avoided-infections.pdf}
    \caption{\textbf{Analysis of epidemiological data and estimated relative risk of infection with air cleaners.}. \textbf{(a)}~Weekly number of respiratory cases with and without air cleaners. \textbf{(b)}~Estimated mean number of infections (posterior mean as dot and 50\%, 80\%, and 95\%-CrI as shaded areas) if, in both classes, air cleaners had been installed throughout the study or not been installed.}
    \label{fig:comparison}
\end{figure}

@@

We detected few respiratory viruses in bioaerosol samples (3~positive bioaoerosl and 50~positive saliva samples), especially when comparing to our previous study\cite{Banholzer2023PLoSMed} (10~positive bioaerosol and 21~positive saliva samples). The number of viruses detected on the filters of the air cleaners was also lower (this study: 4~positive samples in 160~swabs, previous study: 8~positive samples in 80~swabs). The bioaerosol sampling devices were the same in both studies and technical issues were not observed. It is also unlikely that environmental factors are responsible for the differences. Temperature and humidity were similar, and CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm in this vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Therefore, we think that the differences in the number of positive bioaerosols samples between both studies is most likely related to differences in the distribution of respiratory viruses. Strikingly, we detected a wide range of respiratory viruses (Flu B, HRV, AdV, etc) in this study, as compared to only SARS-CoV-2 in our previous study. The shift in the pattern of respiratory viruses has been observed in other studies\cite{Nygaard2023Lancet,Sauteur2022EuroSurv}. SARS-CoV-2 may be present in higher concentrations in air and may be more easily detected in bioaerosol samples due to differences in physicochemical properties of virus-laden aerosols (\eg physical size, viral load, and infectivity) that influence the generation and survival of airborne viral RNA\cite{Wang2021,Sattar2016Book}. In addition, although respiratory viruses have been detected in bioaerosols (\eg Flu\cite{Bischoff2013JID,Pan2017mSphere}, HRV\cite{Myatt2004AJRCCM,Fabian2011JAMPDD}, AdV\cite{Nguyen2017OFID,Pan2017mSphere}, and CoV\cite{Guo2020EID,Liu2020Nature}), bioaerosol sampling remains a technical challenge\cite{Wang2021 ,Belser2023PLOSPath,Bekking2019IORV,Mainelis2020AST}.

>>>>

In a previous study\cite{Banholzer2023PLoSMed}, we estimated the effectiveness of mask wearing and air cleaners during the SARS-CoV-2 Omicron wave. \Cref{fig:comparison} shows a comparison of the results. The distribution of respiratory infections is different (\Cref{fig:comparison}a) because we almost exclusively detected SARS-CoV-2 in our previous study during the COVID-19 pandemic, as compared to a range of respiratory infections in this study during non-pandemic times. Although there were more positive saliva samples in this study, the number of respiratory infections detected in bioaerosols and on the HEPA-filters of the air cleaners was lower (\Cref{fig:comparison}b).
The reductions in particle concentrations were larger and with lower uncertainty in this study (\Cref{fig:comparison}c). The risk of infection tended to be lower with air cleaners in this study where air cleaners were analyzed using a cross-over design, but not in our previous study where air cleaners were installed only at the end of the study when most students had already been infected with SARS-Cov-2(\Cref{fig:comparison}d).

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/comparison.pdf} 
    \caption{\textbf{Comparison of the results between our previous study in 2022 during the COVID-19 pandemic (blue) and this study in 2023 during non-pandemic times (black).}. \textbf{(a)}~Proportion of positive saliva samples and virus. Viruses: Influenza B (Flu B), rhinovirus (HRV), adenovirus (AdV), SARS-CoV-2 (CoV), metapneumovirus (MPV), parainfluenza virus (PIV). \textbf{(b)}~Number of positive saliva, bioaerosol, and HEPA-filter samples. \textbf{(c)}~Estimated reduction in aerosol number and particle mass concentrations (CN in 1/cm$^3$) and particle mass concentration (PM for particles of sizes <1 to <10~$\mu$m, respectively in $\mu$gm$^{-3}$) with air cleaners (posterior mean as squares and 95\%-CrI as lines). \textbf{(d)}~Adjusted infection risk ratio (posterior mean as squares and 95\%-CrI as lines).}
    \label{fig:comparison}
\end{figure}

@@

We detected few respiratory viruses in bioaerosol samples (3~positive bioaoerosl and 50~positive saliva samples), especially when comparing to our previous study\cite{Banholzer2023PLoSMed} (10~positive bioaerosol and 21~positive saliva samples). The number of viruses detected on the filters of the air cleaners was also lower (this study: 4~positive samples in 160~swabs, previous study: 8~positive samples in 80~swabs). The bioaerosol sampling devices were the same in both studies and technical issues were not observed. It is also unlikely that environmental factors are responsible for the differences. Temperature and humidity were similar, and CO$_2$ levels were even higher in this study (1,702$\pm$370\,ppm in this vs 1,064$\pm$232\,ppm in the previous study), suggesting that classrooms were not as well ventilated as during the COVID-19 pandemic. Lower ventilation was potentially offset by a larger reduction in aerosol and particle concentrations through air cleaners in this study. However, it could be that air cleaners removed particles less quickly than ventilation. The most striking difference is in the distribution of respiratory viruses. Compared to the COVID-19 pandemic, we detected a wide range of respiratory viruses (Flu B, HRV, AdV, etc) in this study. The shift in the pattern of respiratory viruses has been observed in other studies\cite{Nygaard2023Lancet,Sauteur2022EuroSurv}. SARS-CoV-2 may be present in higher concentrations in air and may be more easily detected in bioaerosol samples due to differences in physicochemical properties of virus-laden aerosols (\eg physical size, viral load, and infectivity) that influence the generation and survival of airborne viral RNA\cite{Wang2021,Sattar2016Book}. In addition, although various respiratory viruses have been detected in bioaerosols (\eg Flu\cite{Bischoff2013JID,Pan2017mSphere}, HRV\cite{Myatt2004AJRCCM,Fabian2011JAMPDD}, AdV\cite{Nguyen2017OFID,Pan2017mSphere}, and CoV\cite{Guo2020EID,Liu2020Nature}), bioaerosol sampling remains a technical challenge\cite{Wang2021 ,Belser2023PLOSPath,Bekking2019IORV,Mainelis2020AST}.
